---
output: pdf_document
---
## Co-Occurrence of Words in the Shakespeare Drama Corpus
***
The following co-occurrence script aims to discover the semantic proximity of two words throughout the Shakespeare Drama Corpus. At the end, it will take in a word of the user's choice and find the top ten closest terms by proximity. [ SOMETHING ABOUT USING A LOT OF FUNCTIONS FROM PACKAGES ]

***

### Global parameters 
***
Set working directory by pointing to the location on your computer where you have stored the files. Below, we have chosen to Save the folder "RAnalysis" on the Desktop on a Mac. It contains all the other R scripts, texts, notebooks, and results. If you have branched the Github, simply note where you have save the folder. If you are on a PC, you will need to use an absolute path such as "C:Users:XXX."

Hint: Can't figure out your syntax? Click Session > Set Working Directory > Choose Directory, then select the Text_Analysis directory in which you are working. This will set your working directory in the console below while you are working here, but make sure to copy the path into the "setwd" command below to keep the directory constant if you close this script and reopen later.
***
```{r}
setwd("~/Documents/IU/CyberDH/R/Text_Analysis/RNotebooks")
```

***

#### Include necessary packages for notebook 

R's extensibility comes in large part from packages. Packages are groups of functions, data, and algorithms that allow users to easily carry out processes without recreating the wheel. Some packages are included in the basic installation of R, others created by R users are available for download. Make sure to have the following packages installed before beginning so that they can be accessed while running the scripts. By calling library(packagename) they will be automatically loaded and ready to use.

The first three packages are used to render the RNotebook you are currently viewing:

knitr - Creates a formatted report from the script provided

markdown - a package used to render textual documents from plain text to others such as R and XML

rmarkdown - similar to markdown but specifically to render R documents


The next two packages are used within the co-occurrence script:

tm - this package provides tools (functions) for performing various types of text mining. In this script, we will use tm to performing text cleaning in order to have uniform data for analysis. Check out [this link](https://cran.r-project.org/web/packages/tm/vignettes/tm.pdf) for the documentation!

RWeka - another package providing tools for 


```{r}
library(knitr)
library(markdown)
library(rmarkdown)
library(tm)
library(RWeka)
```

***