---
output: pdf_document
---

***
### Beginning Text Preparation

In order to perform text analysis, there are a few R commands you should have up your sleeve. Some of the commands help get you set up and locate all of the files in your corpora. Other commands can be used throughout the programming process to check on your algorithm and make sure everything looks the way you think it should. Learning the following commands will give you a brief introduction to R while also setting you up with a solid toolkit to begin programming.

***
### Loading files from Github to Karst

Since we will be using Karst as the example for our file paths and other elements in these R notebooks, many of you may want to save these files to Karst to make following along a bit easier. To do this you will need a Karst account if you do not already have one. Indiana University students, faculty, staff, and sponsored affiliates can request a Karst account. Steps on how to do so can be found [here](https://kb.iu.edu/d/bezu#account). Once you have your account, you'll want to download the Text-Analysis folder and then upload it to Karst. Information on how to access your Karst account can be found [here](https://kb.iu.edu/d/bezu#access).



***
### Global parameters

In your terminal or command prompt, launch R by typing "R" (without quotes) and pressing Enter. Next, set the working directory by pointing to the location on your computer where you have stored the files. Below, we have chosen to Save the folder "Text-Analysis" in the Cyber DH account on the Karst super-computer here at Indiana University. It contains the R scripts, texts, notebooks, and results. If you have cloned the Github repository, simply point to where you have saved the folder. If you save it to your personal Karst folder, it will most likely look very similar to the example below. You start in the "N" or "gpfs" directory, either one will work and the default directory Karst starts you in seems to vary by user. We will be using "N" in our examples. From here you go to the "home" directory, and this is true for "N" or "gpfs." Now you find the first letter of your username, and using cyberdh as an example, we go to "c" (and yes the case of the letters matter as "N" is different than "n"). Then we look for the second letter in the username, which is "y" for our purposes here. Then you find your full username, which for us is "cyberdh" and then you should see a folder named "Karst". This is where the similarities might end as it now depends on where you saved your "Text-Analysis" folder. The final path should look similar to this: "/N/home/x/x/xxxxxxx/Karst/Text-Analysis" (with the quotes now). Alternatively, if you are on a PC, you will need to use an absolute path such as "C:/Users/XXX" (with the quotes again). If you decide to save it to a Mac you can use a relative path such as "~/Desktop/XXX" (also with the quotes again).

OR

If you are using RStudio, click Session in the menu bar > Set Working Directory > Choose Directory, then select the Text-Analysis directory in which you are working. This will set your working directory in the console pane, but make sure to copy the path into the source pane above to keep the directory constant if you close this script and reopen later.

HINT: Your working directory is the folder from which you will be pulling your texts.

```{r}
setwd("N/home/c/y/cyberdh/Text-Analysis/data")
```

***
### Include necessary packages for notebook 

R's extensibility comes in large part from packages. Packages are groups of functions, data, and algorithms that allow users to easily carry out processes without recreating the wheel. Some packages are included in the basic installation of R, others created by R users are available for download. Make sure to have the following packages installed before beginning so that they can be accessed while running the scripts.

OR

If you are using RStudio, packages can be installed by navigating to Tools in the menu bar > Install Packages.

```{r}
library(knitr)
```

***
### Load Packages

In order to access the packages you have installed within the environment in which you are currently working, you must load them at the beginning of your script. To load packages, use the library() command (see code below).

The three packages listed are used to render the RNotebook you are currently viewing:

knitr - Creates a formatted report from the script provided

markdown - a package used to render textual documents from plain text to others such as R and XML

rmarkdown - similar to markdown but specifically to render R documents

```{r}
library(knitr)
library(markdown)
library(rmarkdown)
```

To perform topic modeling, you may load MALLET or tm. For data visualization, a popular package is ggplot2, or perhaps hclust to create a cluster dendrogram. You can peruse the various contributed packages [here](https://cran.r-project.org/web/packages/available_packages_by_name.html). The tutorials included in the R Toolkit will instruct you which packages to install and load.

***
### Load data 

Now you are ready to start looking at your data! First, you must load it into your environment. The scan() function will do this for you. If you want to load just one text into your environment, here is the syntax:

The first argument is the filename (or path if the file resides in a different directory than your working directory). 
The second argument "what" specified as type "character" will read the text in as a character vector. 
The third argument "sep" specified as "backslash" + "n" which is the way to code line breaks in R.

So putting everything together, this line reads in JaneEyre.txt, and separates the text into a character vector by line.

```{r}
text <- scan("/N/home/c/y/cyberdh/Text-Analysis/data/bronte/janeEyre.txt", what="character", sep="\n")
```

The final, crucial aspect of this line is the assignment. "<-" assigns whatever results on the right side of the arrow into the variable specified on the left side. Some programming languages use "=" instead of the arrow. R will also acknowledge this, but using the arrow is best practice.

Here, we have named that variable "text" since it holds the text with which we are working. However, you can name this variable whatever you would like. This line will give the exact same result, although it is best to name the variable in relation to what it holds:

```{r}
potatoes <- scan("/N/home/c/y/cyberdh/Text-Analysis/data/bronte/janeEyre.txt", what="character", sep="\n")
```

Now that we have the text saved as a variable, we can reuse that variable simply by calling "text" instead of the entire scan line again. The next few commands will use "text" to explore the data.

***
### R Objects

R is distinct from other programming languages in that it handles objects a little differently. Throughout text analysis, you will need to massage your text and textual data by changing it into various kinds of objects which make things easier. Check out [this tutorial](https://www.nealgroothuis.name/introduction-to-data-types-and-objects-in-r/) by Neal Groothuis which simply explains the various types of data objects. Since some kinds of objects prohibit certain actions, a simple way to check the type of object you are currently working with is class(). For example, the code below shows that the class of the Jane Eyre "text" is a character vector, which we would assume since that is what we specified while loading it in.

```{r}
class(text)
```

***
### Data Inspection

Just as you may want to verify the type of the object with which you are working, you may want to view it from other angles to make sure the data is formed as you expect.

The length function shows the number of elements within an object. If you find that the length is zero, you may have to go back and reload the data, or check to make sure your algorithm is working correctly.

```{r}
length(text)
```

We can also look at individual elements - Lets see what the first line of our text is...

```{r}
text[1]
```

Or perhaps you would like to see the first few elements:

```{r}
head(text)
```

Or last few elements:

```{r}
tail(text)
```

There are many more ways to inspect parts of data (check out the CRAN), but these quick checks are helpful while manipulating the data and debugging the inevitable issues you will encounter while developing your script.

***
### Explore

The above commands are a few tips and tricks to get you started with R. Similar to R's extensibility with packages, the R user community has great resouces for learners. The [CRAN FAQ](https://cran.r-project.org/) and the [CRAN Manual](https://cran.r-project.org/doc/manuals/R-intro.pdf) answers quite a few questions about R and its uses.

Googling the issue, function, or object name with "r" will return helpful resources. If a PDF from cran.r-project.org appears, there you will find extensive documentation and examples for that function, etc. and other related resources. Similarly, any result from r-bloggers.com will most likely be helpful. For any other issues, Stack Overflow is helpful to find answers to common questions as well as ask your own.

The rest of the IU tutorials explain some methods for textual analysis using R. If you are ready to dive in, click on one to begin!
