---
output: pdf_document
---
***
## Wordcloud of Top 75 Words in Charlotte Bronte's Jane Eyre

Using the text of Jane Eyre from Project Gutenberg, we will plot the top 75 words in the novel. [http://www.gutenberg.org/cache/epub/1260/pg1260.txt]

***
### Global parameters 

In your terminal or command prompt, launch R by typing "R" (without quotes) and pressing Enter. Next, set the working directory by pointing to the location on your computer where you have stored the files. Below, we have chosen to Save the folder "Text-Analysis" on the Desktop of a Mac as an example. It contains the R scripts, texts, notebooks, and results. If you have cloned the Github repository, simply point to where you have saved the folder. If you are on a PC, you will need to use an absolute path such as "C:/Users/XXX" (with the quotes this time!).

OR

If you are using RStudio, click Session in the menu bar > Set Working Directory > Choose Directory, then select the Text-Analysis directory in which you are working. This will set your working directory in the console pane, but make sure to copy the path into the source pane above to keep the directory constant if you close this script and reopen later.

```{r}
setwd("~/Desktop/Text-Analysis/RNotebooks/")
```

***
#### Include necessary packages for notebook 

R's extensibility comes in large part from packages. Packages are groups of functions, data, and algorithms that allow users to easily carry out processes without recreating the wheel. Some packages are included in the basic installation of R, others created by R users are available for download. Make sure to have the following packages installed before beginning so that they can be accessed while running the scripts.

OR

If you are using RStudio, packages can be installed by navigating to Tools in the menu bar > Install Packages.

```{r}
library(knitr)
library(markdown)
library(rmarkdown)
library(wordcloud)
library(tm)
library(RColorBrewer)
```

Load data (plaintext file), telling the computer that it is formatted by line breaks  

```{r}
text_raw<-scan("~/Desktop/Text-Analysis/data/bronte/janeEyre.txt", what="character", sep="\n")
```

***
####Prepare text data

The tm library works with a "corpus," which can represent a single text, many texts, and even different types of inputs, such as CSV files. In this case, we only have one text in our corpus, but will follow tm's naming conventions for the variables.

```{r}
#Create a corpus 
corpus <- Corpus(VectorSource(text_raw))
```

One of the great things about using tm is that it has many text cleaning functions built into it. Here we have removed extra white space, transformed all the text to lower case, removed stopwords, and removed punctuation. It is possible to simply use an "english" (or other language) stopword list, but I have found the SMART list to be more complete. For example, "said" is on the SMART stopword list but not on the standard "english" list. 

```{r}
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removeWords, stopwords("SMART"))
corpus <- tm_map(corpus, removePunctuation)
```

Plot the wordcloud! There are a number of arguments you can customize: "random.order" is false so that words are plotted in order of decreasing frequency; "scale"" indicates the size of the words; "rot.per" lets you customize the proportion of words that are rotated 90 degrees; "max.words" controls how many words show up in the wordcloud; and we've used the library "RColorBrewer" to give us access to some predefined palettes. Note that if you change the palette, you need to tell the argument how many colors are in the new palette. 

```{r}
wordcloud(corpus,random.order=FALSE,scale=c(3,.1),rot.per=0,
          max.words=75,colors=brewer.pal(8, "Dark2"))
```

***
### Voila!
